---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Getting Set Up

We need the latest versions of these R packages installed from GitHub. 

It's best to run each line in turn in case there are issues or the user needs to provide input. 

```{r setup, eval = FALSE}
## make sure we have the packages we need

if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
req1 <- c("dplyr", "ggplot2")
remotes::install_cran(req1)
## and some github packages
req2 <- c("ropensci/antanym", "AustralianAntarcticDivision/blueant",
         "AustralianAntarcticDivision/raadtools",
         "AustralianAntarcticDivision/SOmap")
remotes::install_github(req2)

```

We can get everything we need for the workshop by running the following code. 

```{r get-data, eval = FALSE}

td <- "."  ## Mike, change to tempdir()
td <- normalizePath("~")
library(dplyr)
my_data_dir <- file.path(td, "data-library")
dir.create(my_data_dir, showWarnings = FALSE)
library(blueant)
src <- bind_rows(
  sources("NSIDC SMMR-SSM/I Nasateam sea ice concentration", hemisphere = "south",
          time_resolutions = "day", years = 2013),
  sources("ETOPO2 bathymetry"))
result <- bb_get(src, local_file_root = my_data_dir, clobber = 0, verbose = TRUE,
                 confirm = NULL)

raadfiles::set_raad_data_roots(my_data_dir, replace_existing = TRUE)
raadtools::set_data_roots(my_data_dir)

```


## Undecided section

This workshop may include a pre-bundled data set using the `piggyback` GitHub releases mechanism. The bowerbird package will happily build this project from scratch from
the original data providers, but it takes longer than we have for this workshop. 

```{r pigback, eval = FALSE, echo = FALSE}
## file nesting doesn't work well here
#system("tar zcvf datalibrary.tar.gz data-library"))
piggyback::pb_upload("data-library/datalibrary.tar.gz", tag = "v0.1.0", name = "datalibrary.tar.gz")
piggyback::pb_download("datalibrary.tar.gz", dest = "data-library/datalibrary.tar.gz" )
```
